---
title: "Vancouver Weather"
author: ""
date: "`r Sys.Date()`"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    includes:
      in_header: gtm-head-jyuill-github-io.html
      before_body: gtm-body-jyuill-github-io.html
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(lubridate)
library(scales)
library(plotly)

theme_set(theme_light())
```

```{r}
## Get data
## get latest data using van-weather-import.R
## import saved data
vw_data <- read_csv("output/van-weather.csv") 

## convert Year, Month to factors for better charting
vw_data$Year <- as.factor(vw_data$Year)
vw_data$Month <- as.factor(vw_data$Month)
vw_data$Season <- as.factor(vw_data$Season)
## set Season order
vw_data$Season <- factor(vw_data$Season, levels=c("Winter","Spring","Summer","Fall"))
#levels(vw_data$Season)

```

```{r}
## Monthly precipitation for each month, each year
vw.precip.mth.yr <- vw_data %>% mutate(precip.day=ifelse(Total.Precip>0,1,0)) %>% 
  group_by(Year, Month) %>%
  summarize(ttl_precip=sum(Total.Precip, na.rm=TRUE),
            precip_days=sum(precip.day, na.rm=TRUE))

## Group by month
vw.precip.mth <- vw.precip.mth.yr %>% group_by(Month) %>%
  summarize(ave_precip=mean(ttl_precip, na.rm=TRUE)) %>%
  mutate(pc_rank=percent_rank(ave_precip),
         pc=ave_precip/sum(ave_precip),
         rank=rank(desc(ave_precip)))

## filter for current year to overlay current year metrics on chart
vw.precip.mth.curr.yr <- vw.precip.mth.yr %>% filter(Year==year(Sys.Date()))
## need to have full 12 months for chart (even though some empty)
## get highest month with data in current year
fms <- max(as.numeric(vw.precip.mth.curr.yr$Month))+1
## create data frame for blank months
vw.precip.yr.future.mth <- data.frame(Year='2019',
                                      Month=c(fms:12),
                                       ttl_precip=NA,
                                      precip_days=NA)
## convert Month field in current yr from factor to integer for binding
vw.precip.mth.curr.yr$Month <- as.integer(vw.precip.mth.curr.yr$Month)
## add blank rows to current yr
vw.precip.mth.curr.yr <- bind_rows(vw.precip.mth.curr.yr,
                                    vw.precip.yr.future.mth) 
```

Current
===================================================================

Column
-----------------------------------------------------------------------

### Recent Precipitation

```{r}
num_days <- 30
p1 <- vw_data %>% filter(Date>=max(vw_data$Date)-num_days) %>%
  ggplot(aes(x=Date, y=Total.Precip))+geom_col()+
  scale_y_continuous(expand=expansion(add=c(0,2)))+
  labs(x="", y="Total Daily Precipitation")

ggplotly(p1)
```

### Monthly Precipation (red diamonds = current yr)

```{r}
## Ave precip by month ####
# chart.title <- "Average Precipitation by Month"
# sub.title <- "(red marker is curr. yr.)"
# p2 <- ggplot(vw.precip.mth, aes(x=Month, y=ave_precip))+geom_col()+
#   geom_point(data=vw.precip.mth.curr.yr, aes(x=Month, y=ttl_precip),shape=18, color='red', size=2.5)+
#   scale_y_continuous(expand=c(0.1,0.1))+
#   labs(title=chart.title, subtitle=sub.title, x="", y="Ave. Precipitation")
# ggplotly(p2)


## Range in precipitation by month ####
chart.title <- "Monthly Precip. Ranges during Period"
sub.title <- "(red markers: current yr.)"
p2 <- ggplot(vw.precip.mth.yr, aes(x=Month, y=ttl_precip))+geom_boxplot()+
  geom_point(data=vw.precip.mth.curr.yr, aes(x=Month, y=ttl_precip),shape=18, color='red', size=2.5)+
  labs(x="", y="Mthly Precipitation")
ggplotly(p2)

```

Column
-----------------------------------------------------------------------

### Recent Temperature

```{r}
num_days <- 30
t1 <- vw_data %>% filter(Date>=max(vw_data$Date)-num_days) %>%
  ggplot(aes(x=Date, y=Max.Temp))+geom_line(color='red')+
  geom_line(aes(y=Mean.Temp), color='green')+
  geom_line(aes(y=Min.Temp), color='blue')+
  #scale_y_continuous(expand=expansion(add=c(0,2)))+
  scale_x_date(expand=c(0,0))+
  labs(x="", y="Max Daily Temperature")

ggplotly(t1)
```

### Monthly Temperature Comp. (red diamonds = current yr)

```{r}
## Monthly precipitation for each month, each year
vw.temp.mth.yr <- vw_data %>% 
  group_by(Year, Month) %>%
  summarize(ave_temp=mean(Mean.Temp, na.rm=TRUE),
            ave_max_temp=mean(Max.Temp, na.rm=TRUE),
            ave_min_temp=mean(Min.Temp, na.rm=TRUE),
            max_temp=max(Max.Temp, na.rm=TRUE),
            min_temp=min(Min.Temp, na.rm=TRUE))

#is.nan(vw.temp.mth.yr$ave_max_temp)
#vw.temp.mth.yr %>% filter(max_temp==-Inf)
## need to clean up missing values
## even with na.rm used above, if there are only NA for a given month, 
##  it will return NaN for mean and Inf or -Inf for min/max
## clean up
vw.temp.mth.yr <- vw.temp.mth.yr %>% mutate(
  ave_temp=ifelse(is.nan(ave_temp),NA,ave_temp),
  ave_max_temp=ifelse(is.nan(ave_max_temp),NA,ave_max_temp),
  ave_min_temp=ifelse(is.nan(ave_min_temp),NA,ave_min_temp),
  max_temp=ifelse(max_temp==-Inf|max_temp==Inf,NA, max_temp),
  min_temp=ifelse(min_temp==-Inf|min_temp==Inf,NA, min_temp)
)

## Monthly average temp across years
vw.temp.mth <- vw.temp.mth.yr %>% group_by(Month) %>%
  summarize(ave_temp=mean(ave_temp, na.rm=TRUE),
            ave_max_temp=mean(ave_max_temp, na.rm=TRUE),
            ave_min_temp=mean(ave_min_temp, na.rm=TRUE)) %>%
  mutate(pc_rank=percent_rank(ave_temp),
         rank=rank(desc(ave_temp)))

## filter for current year to overlay current year metrics on chart
vw.temp.mth.curr.yr <- vw.temp.mth.yr %>% filter(Year==year(Sys.Date()))
## need to have full 12 months for chart (even though some empty)
## get highest month with data in current year
fms <- max(as.numeric(vw.temp.mth.curr.yr$Month))+1
## create data frame for blank months
vw.temp.yr.future.mth <- data.frame(Year=year(Sys.Date()),
                                      Month=c(fms:12))
## convert Month field in current yr from factor to integer for binding
vw.temp.mth.curr.yr$Year <- as.integer(vw.temp.mth.curr.yr$Year)
vw.temp.mth.curr.yr$Month <- as.integer(vw.temp.mth.curr.yr$Month)

## add blank rows to current yr
vw.temp.mth.curr.yr <- bind_rows(vw.temp.mth.curr.yr,
                                    vw.temp.yr.future.mth) 

```

```{r}
## Distribution of temperature by month ####

## Range in temperature by month ####
chart.title <- "Monthly Temp. Ranges during Period (Red markers: current yr.)"
t2 <- ggplot(vw.temp.mth.yr, aes(x=Month, y=ave_temp))+geom_boxplot()+
  geom_point(data=vw.temp.mth.curr.yr, aes(x=Month, y=ave_temp),shape=18, color='red', size=2.5)+
  labs(x="", y="Ave. Temperature") ## subtitle doesn't work with ggplotly
ggplotly(t2)

```


Source
===================================================================

Produced by John Yuill <br />
All source code and data used is available in my Github repo: <br />
[proj-r-van-weather](https://github.com/jyuill/proj-r-van-weather)

**Govt of Canada Historical Weather data** <br />
Data is collected from: <br />
http://climate.weather.gc.ca/historical_data/search_historic_data_e.html

Data is available from the government on a daily basis and updated *occasionally* here. ;) 

Date range currently covered:

**Earliest date:** *`r min(vw_data$Date)`* <br />
**Most recent date:** *`r max(vw_data$Date)`* <br />

The following weather stations are represented. Not the preferred approach but I couldn't find continuous data for any single weather station.

```{r, fig.height=1.5, fig.width=7}
## stations
vw_stn <- vw_data %>% select(Station, Year) %>% mutate(
  Year=as.numeric(levels(vw_data$Year))[Year]
) 

vw_stn %>% ggplot(aes(x=Station, y=Year))+geom_point()+
  coord_flip()+theme_classic()+
  labs(title='Weather Stations',x='')

```

### Data Structure

Check data structure and summary:

```{r DATA STRUCTURE}
str(vw_data)

summary(vw_data)
```
