---
title: "Vancouver Weather Analysis"
output:
  html_document:
    df_print: paged
---

```{r SETUP, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, 
                      warning=FALSE, 
                      message=FALSE)

library(tidyverse)
library(scales)
library(PerformanceAnalytics)
library(lubridate)
library(plotly)

## chart colors
color.line <- 'grey70'

```

# Intro {.tabset}

Basic analysis of Vancouver weather patterns:

* from [Govt. of Canada Historical Weather](http://climate.weather.gc.ca/historical_data/search_historic_data_e.html) Data website
* monthly
* precipitation, temperature

## Data Collection

Data is collected from Govt of Canada website:

1. search: http://climate.weather.gc.ca/historical_data/search_historic_data_e.html
2. select station of interest by searching for stations with 'vancouver' at start of name 
3. select 'daily', select 'year' (note that year shown is end of period data is available for; most stations have very little data)
4. 'Go' goes to a page to access results for year

(process is automated in 'van-weather-import.R')

```{r IMPORT DATA}
## get latest data using van-weather-import.R
## import saved data
vw_data <- read_csv("output/van-weather.csv") 

## convert Year, Month to factors for better charting
vw_data$Year <- as.factor(vw_data$Year)
vw_data$Month <- as.factor(vw_data$Month)
vw_data$Season <- as.factor(vw_data$Season)
## set Season order
vw_data$Season <- factor(vw_data$Season, levels=c("Winter","Spring","Summer","Fall"))
levels(vw_data$Season)

```

**Earliest date:** *`r min(vw_data$Date)`* <br />
**Most recent date:** *`r max(vw_data$Date)`* <br />
`
```{r, fig.height=1.5, fig.width=7}
## stations
vw_stn <- vw_data %>% select(Station, Year) %>% mutate(
  Year=as.numeric(levels(vw_data$Year))[Year]
) 

vw_stn %>% ggplot(aes(x=Station, y=Year))+geom_point()+
  coord_flip()+theme_classic()+
  labs(title='Weather Stations',x='')

```

Check data structure and summary:

```{r DATA STRUCTURE}
str(vw_data)

summary(vw_data)
```

* Looks like quite a few NAs for temp and precip; as long as they are spread out, shouldn't be a problem

Check Data Relationships

* uses PerformanceAnalytics package

```{r DATA RELATIONSHIPS}
chart.Correlation(vw_data[,c(6:9)], histogram = TRUE)

```

* Looks like temperatures are highly correlated and precipation has weak negative correlation with temperate (colder weather slightly more associated with rain)

```{r TS1}
## time series chart from Performance Analytics pkg - not really better than just using ggplot2
# vw_ts <- data.frame(vw_data)
# row.names(vw_ts) <- vw_ts$Date
# chart.TimeSeries(vw_ts[,c(5,6,7,8), drop=FALSE])
```

## Precipitation 

### Monthly precipitation data: Averages

```{r MONTHLY PRECIP}
## Monthly precipitation for each month, each year
vw.precip.mth.yr <- vw_data %>% mutate(precip.day=ifelse(Total.Precip>0,1,0)) %>% 
  group_by(Year, Month) %>%
  summarize(ttl_precip=sum(Total.Precip, na.rm=TRUE),
            precip_days=sum(precip.day, na.rm=TRUE))
```


```{r MONTHLY PRECIP COMBINED}
vw.precip.mth <- vw.precip.mth.yr %>% group_by(Month) %>%
  summarize(ave_precip=mean(ttl_precip, na.rm=TRUE)) %>%
  mutate(pc_rank=percent_rank(ave_precip),
         pc=ave_precip/sum(ave_precip),
         rank=rank(desc(ave_precip)))

## filter for current year to overlay current year metrics on chart
vw.precip.mth.curr.yr <- vw.precip.mth.yr %>% filter(Year==year(Sys.Date()))
## need to have full 12 months for chart (even though some empty)
## get highest month with data in current year
fms <- max(as.numeric(vw.precip.mth.curr.yr$Month))+1
## create data frame for blank months
vw.precip.yr.future.mth <- data.frame(Year='2019',
                                      Month=c(fms:12),
                                       ttl_precip=NA,
                                      precip_days=NA)
## convert Month field in current yr from factor to integer for binding
vw.precip.mth.curr.yr$Month <- as.integer(vw.precip.mth.curr.yr$Month)
## add blank rows to current yr
vw.precip.mth.curr.yr <- bind_rows(vw.precip.mth.curr.yr,
                                    vw.precip.yr.future.mth) 

## Ave precip by month ####
chart.title <- "Average Precipitation by Month (red marker is curr. yr.)"
ggplot(vw.precip.mth, aes(x=Month, y=ave_precip))+geom_col()+
  geom_point(data=vw.precip.mth.curr.yr, aes(x=Month, y=ttl_precip),shape=18, color='red', size=5)+
  scale_y_continuous(expand=c(0,0))+
  ggtitle(chart.title)+theme_classic()
## Ave precip by month in order
chart.title <- "Months Ranked by Ave. Precipitation"
ggplot(vw.precip.mth, aes(x=reorder(Month, -ave_precip), y=ave_precip))+geom_col()+
  geom_hline(yintercept = mean(vw.precip.mth$ave_precip), linetype='dashed', color=color.line)+
  labs(title=chart.title, subtitle = '(Dashed line is average)')+
  theme_classic()

```

#### Data by Month - ranked in order of highest precipitation

```{r MONTHLY PRECIP TABLE}
vw.precip.mth.rank <- vw.precip.mth %>% arrange(rank)
vw.precip.mth.rank
```

### Monthly Preciptiation: Ranges

```{r PRECIP MTH BOX}
## Distribution of precipitation by month ####

## Range in precipitation by month ####
chart.title <- "Monthly Precip. Ranges during Period (Red markers: current yr.)"
ggplot(vw.precip.mth.yr, aes(x=Month, y=ttl_precip))+geom_boxplot()+
  geom_point(data=vw.precip.mth.curr.yr, aes(x=Month, y=ttl_precip),shape=18, color='red', size=5)+
  ggtitle(chart.title)+theme_classic()
## Range in precip days by month
chart.title <- "Monthly Rainy Days Ranges during Period (Red markers: current yr.)"
ggplot(vw.precip.mth.yr, aes(x=Month, y=precip_days))+geom_boxplot()+
  geom_point(data=vw.precip.mth.curr.yr, aes(x=Month, y=precip_days),shape=18, color='red', size=5)+
  ggtitle(chart.title)+theme_classic()
```


### Current Month Precipitation

```{r CURRENT MTH Precip}
## get name of current month
curr.mth <- month.name[month(Sys.Date())]

## filter monthly data set for current month only
vw.precip.mth.curr <- vw.precip.mth.yr %>% filter(Month==month(Sys.Date())) %>%
  arrange(desc(Year)) ## sort by year to get latest

## get recent years for current month precipitation
recentXyr <- 12
vw.precip.mth.curr.recent <- vw.precip.mth.curr[1:recentXyr,]
chart.title <- paste0(curr.mth, " Rain for each Year (last ",recentXyr, " years - with ave.)")
ggplot(vw.precip.mth.curr.recent, aes(x=Year, y=ttl_precip))+geom_bar(stat='identity')+
  geom_hline(yintercept=mean(vw.precip.mth.curr.recent$ttl_precip), linetype='dashed', color=color.line)+
  scale_y_continuous(expand=c(0,0))+theme_classic()+
  ggtitle(chart.title)

## show years in order of precipitation for current month
vw.precip.mth.curr <- vw.precip.mth.yr %>% filter(Month==month(Sys.Date())) %>%
  arrange(desc(ttl_precip)) ## sort by precipitation
topXyr <- 15
vw.precip.mth.curr.top <- vw.precip.mth.curr[1:topXyr,]

chart.title <- paste0(curr.mth, " Ranked by Precipitation (top ",topXyr," yrs)")
ggplot(vw.precip.mth.curr.top, aes(x=reorder(Year, -ttl_precip), y=ttl_precip))+geom_col()+
  geom_hline(yintercept = mean(vw.precip.mth.curr.top$ttl_precip), linetype='dashed', color=color.line)+
  scale_y_continuous(expand=c(0,0))+theme_classic()+
  labs(title=chart.title, subtitle='(Line is average)')

## get percentiles for current mth
vw.precip.mth.curr.pc <- vw.precip.mth.curr %>% ungroup() %>% arrange(ttl_precip) %>%
  #mutate(precip.pc=quantile(ttl_precip))
  mutate(precip.rank=rank(ttl_precip),
         precip.pc=precip.rank/max(precip.rank),
         cum.dist=cume_dist(ttl_precip),
         pc.rank=percent_rank(ttl_precip))

```

### Monthly Comparison YoY

For each month, what has been the pattern in precipitation over the years?

```{r PLOT MONTHS YOY, fig.width=5, fig.height=14}
## show each mth side-by-side YoY to see patterns, highs, lows

## facet_wrap is condensed but too many years to be able to read x axis
# ggplot(vw.precip.mth.yr, aes(x=Year, y=ttl_precip))+geom_bar(stat='identity')+
#   facet_wrap(Month~.)+
#   theme(axis.text.x  = element_text(angle=90, vjust=0.5))

## facet_grid is too wide and short unless re-sized
ggplot(vw.precip.mth.yr, aes(x=Year, y=ttl_precip))+geom_bar(stat='identity')+
  geom_hline(data=vw.precip.mth, aes(yintercept=ave_precip), linetype='dashed', color=color.line)+
  facet_grid(Month~.)+
  theme_bw()+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5))
```

### Seasonal Precipitation

#### Season definitions:

* Winter: Dec, Jan, Feb
* Spring: Mar, Apr, May
* Summer: Jun, Jul, Aug
* Fall: Sep, Oct, Nov

*Note: the year for a winter season is applied to year at end of season. So winter from Dec 2018 to Feb 2019 is considered winter of 2019.*

```{r SEASONAL DATA}
## Monthly precipitation for each season, each year
vw.precip.seas.yr <- vw_data %>% mutate(precip.day=ifelse(Total.Precip>0,1,0)) %>% 
  group_by(Season.Yr, Season) %>%
  summarize(ttl_precip=sum(Total.Precip, na.rm=TRUE),
            precip_days=sum(precip.day, na.rm=TRUE))

## Monthly summary data for each season
vw.precip.seas <- vw.precip.seas.yr %>% group_by(Season) %>%
  summarize(mean.precip=mean(ttl_precip),
          median.precip=median(ttl_precip),
          mean_precip_days=mean(precip_days),
          median.precip=median(precip_days))

## get current year seasonal data for comparison via dots
vw.precip.seas.yr.current <- vw.precip.seas.yr %>% filter(
 Season.Yr==year(Sys.Date())
)

```

```{r PRECIP SEASON BOX}
## Distribution of precipitation by season ####

## Range in precipitation by season ####
chart.title <- "Seasonal Precip. Ranges during Period"
ggplot(vw.precip.seas.yr, aes(x=Season, y=ttl_precip))+geom_boxplot()+
  geom_point(data=vw.precip.seas.yr.current, aes(x=Season, y=ttl_precip),shape=18, color='red', size=5)+
  ggtitle(chart.title)+theme_classic()

```

```{r PLOT SEASON YOY, fig.width=5, fig.height=8}
## show each season side-by-side YoY to see patterns, highs, lows

## facet_wrap is condensed but too many years to be able to read x axis
# ggplot(vw.precip.mth.yr, aes(x=Year, y=ttl_precip))+geom_bar(stat='identity')+
#   facet_wrap(Month~.)+
#   theme(axis.text.x  = element_text(angle=90, vjust=0.5))

## facet_grid is too wide and short unless re-sized
ggplot(vw.precip.seas.yr, aes(x=Season.Yr, y=ttl_precip))+geom_bar(stat='identity')+
  geom_hline(data=vw.precip.seas, aes(yintercept=mean.precip), linetype='dashed', color=color.line)+
  facet_grid(Season~.)+
  theme_bw()+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5))
```

### Seasonal Precipitation Annual Ranking

* Need separate charts because each season ranked individually
* Line = median

```{r PLOT SEASONS RANKED, fig.height=3}
## filter for each season and plot ranked years
## use facet as convenient way to label the chart with season
## reference line for median
vw.precip.s <- vw.precip.seas.yr %>% filter(Season=='Winter')
ggplot(vw.precip.s, aes(x=reorder(Season.Yr,-ttl_precip), y=ttl_precip))+
  geom_bar(stat='identity')+
  geom_hline(yintercept=median(vw.precip.s$ttl_precip), linetype='dashed', color=color.line)+
  facet_grid(.~Season)+
  theme_bw()+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5))

# vw.precip.seas.yr %>% filter(Season=='Spring') %>%
# ggplot(aes(x=reorder(Season.Yr,-ttl_precip), y=ttl_precip))+
#   geom_bar(stat='identity')+
#   facet_grid(.~Season)+
#   theme_bw()+
#   theme(axis.text.x  = element_text(angle=90, vjust=0.5))

vw.precip.s <- vw.precip.seas.yr %>% filter(Season=='Spring')
ggplot(vw.precip.s, aes(x=reorder(Season.Yr,-ttl_precip), y=ttl_precip))+
  geom_bar(stat='identity')+
  geom_hline(yintercept=median(vw.precip.s$ttl_precip), linetype='dashed', color=color.line)+
  facet_grid(.~Season)+
  theme_bw()+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5))

vw.precip.s <- vw.precip.seas.yr %>% filter(Season=='Summer')
ggplot(vw.precip.s, aes(x=reorder(Season.Yr,-ttl_precip), y=ttl_precip))+
  geom_bar(stat='identity')+
  geom_hline(yintercept=median(vw.precip.s$ttl_precip), linetype='dashed', color=color.line)+
  facet_grid(.~Season)+
  theme_bw()+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5))

vw.precip.s <- vw.precip.seas.yr %>% filter(Season=='Fall')
ggplot(vw.precip.s, aes(x=reorder(Season.Yr,-ttl_precip), y=ttl_precip))+
  geom_bar(stat='identity')+
  geom_hline(yintercept=median(vw.precip.s$ttl_precip), linetype='dashed', color=color.line)+
  facet_grid(.~Season)+
  theme_bw()+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5))

```

## Temperature

### Check for missing data

```{r, fig.height=3, fig.width=7}
vw_na <- vw_data %>% filter(is.na(Max.Temp)|is.na(Min.Temp)|is.na(Mean.Temp))
vw_na %>% group_by(Year) %>% summarize(NAs=n()) %>% filter(NAs>1) %>% ggplot(aes(x=reorder(Year, NAs), y=NAs))+
  geom_col()+coord_flip()+theme_classic()+labs(title='Temperature NAs by Yr', x="")

## can also check missing/non-missing by years with table
#table(vw_data$Year, is.na(vw_data$Mean.Temp))

```

* Something to watch for if there are suspicious results for years with higher NAs.
* Probbaly not going to be enough cause to throw whole year out. Could do some imputation of missing values, such as using averages, but probably not worth it.

### Ave. Annual Temperature

```{r ANNUAL TEMP}
## get ave of mean, min, max daily temp by year
## first remove data from current yr if less than at least 9 mths datain yr
if(month(max(vw_data$Date))<10){
  vw.temp.yr <- vw_data %>% filter(as.numeric(as.character(Year))<max(as.numeric(as.character(vw_data$Year))))
} else {
  vw.temp.yr <- vw_data
}
## group by yr to get mean temps
vw.temp.yr <- vw.temp.yr %>% group_by(Year) %>% summarize(mean.mean=mean(Mean.Temp, na.rm=TRUE),                                 mean.min=mean(Min.Temp, na.rm=TRUE), 
                                                       mean.max=mean(Max.Temp, na.rm=TRUE))

## line chart - note need group=1 so ggplot knows how to connect data points for line
chart.title <- 'Ave. Daily Temp by Year, with Trend'
ggplot(vw.temp.yr, aes(x=Year, y=mean.mean, group=1))+geom_line()+
  geom_smooth(method='lm')+theme_classic()+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5))+
  labs(title=chart.title, y='Ave. Daily Temp', x='')

```

* 1985 looks like an outlier, although doesn't look like any NAs based on info above. 
* Overall, appears to be a legitimate trend in rising temperatures

```{r FILTER YRS}
## BASED ON EARLIER DATASET FROM VAN HARBOUR WHICH MUCH MORE MISSING DATA
## NOT NEEDED WITH LATEST DATA
# 
# ## exclude years with missing data - including current yr
# vw.temp.yr.x <- vw.temp.yr %>% filter(Year!=1988 & Year!=1989 &
#                                         Year!=1990 &
#                                           Year!=2007 & 
#                                         Year!=2010 &
#                                         Year!=2011 &
#                                           Year!=year(Sys.Date()))
# 
# ## line chart - note need group=1 so ggplot knows how to connect data points for line
# chart.title <- 'Ave. Daily Temp by Year, with Trend (excl. missing data)'
# ggplot(vw.temp.yr.x, aes(x=Year, y=mean.mean, group=1))+geom_line()+
#   geom_smooth(method='lm')+theme_classic()+
#   theme(axis.text.x  = element_text(angle=90, vjust=0.5))+
#   labs(title=chart.title, y='Ave. Daily Temp', x='')
# 
# ## remove missing years from original data set, since all the code below was
# ## developed in reference to vw.temp.yr
# vw.temp.yr <- vw.temp.yr.x

```

#### Years Ranked by Average Daily Temp

```{r RANK BY YEAR}
## get last yr data
vw.temp.yr.last <- vw.temp.yr %>% filter(Year==year(Sys.Date())-1)

## bar chart - years ranked in order of ave daily temp
chart.title <- 'Years in Order of Ave. Daily Temp. (excl. missing data)'
ggplot(vw.temp.yr, aes(x=reorder(Year, -mean.mean), y=mean.mean))+
  geom_col()+
  #geom_col(data=vw.temp.yr.last, aes(x=Year, y=mean.mean, fill='last yr'))+
  geom_col(data=vw.temp.yr.last, aes(x=Year, y=mean.mean), fill='blue')+
  scale_y_continuous(expand=c(0,0))+
  theme_classic()+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5))+
  labs(title=chart.title,
       subtitle='Most recent yr in blue', y='Ave. Daily Temp', x='')
  
#geom_point(data=vw.precip.mth.curr.yr, aes(x=Month, y=ttl_precip),shape=18, #color='red', size=5)+
```


#### Linear model of annual temperature change

```{r ANNUAL TEMP MODEL}
vw.temp.yr <- vw.temp.yr %>% mutate(
  yr=as.character(Year),
  ynum=as.numeric(yr)
)
at.model <- lm(mean.mean~ynum, vw.temp.yr)

summary(at.model)

slope <- summary(at.model)$coefficients[2,1]
pval <- summary(at.model)$coefficients[2,4]

statsig <- ifelse(pval<0.05,'statistically significant','not statistically significant')
```

Model interpretation: <br />

* Average daily temperature changing `r slope` degrees each yr.
* P-value for statistical significance: `r pval`
* Temperature change is *`r statsig`* over this period

#### Look at Average Daily Maximum and Minimums for each Year

```{r MEAN MIN MAX TEMP}
chart.title <- 'Ave. Daily Temp by Year, with Trend'
ggplot(vw.temp.yr, aes(x=Year, y=mean.mean, group=1))+
  geom_line()+
  geom_line(aes(y=mean.max))+
  geom_line(aes(y=mean.min))+
  geom_smooth(method='lm')+
  theme_classic()+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5))+
  labs(title=chart.title, y='Ave. Daily Temp', x='')
```

Linear modelling for max and min

```{r MAX TEMP MODEL}
at.model <- lm(mean.max~ynum, vw.temp.yr)

summary(at.model)

slope <- summary(at.model)$coefficients[2,1]
pval <- summary(at.model)$coefficients[2,4]

statsig <- ifelse(pval<0.05,'statistically significant','not statistically significant')
```

Model interpretation: <br />

* Average daily MAX temperature changing `r slope` degrees each yr.
* P-value for statistical significance: `r pval`
* Temperature change is *`r statsig`* over this period

```{r MIN TEMP MODEL}
at.model <- lm(mean.min~ynum, vw.temp.yr)

summary(at.model)

slope <- summary(at.model)$coefficients[2,1]
pval <- summary(at.model)$coefficients[2,4]

statsig <- ifelse(pval<0.05,'statistically significant','not statistically significant')
```

Model interpretation: <br />

* Average daily MIN temperature changing `r slope` degrees each yr.
* P-value for statistical significance: `r pval`
* Temperature change is *`r statsig`* over this period

### Monthly average temperature

```{r MONTHLY TEMP}
## Monthly precipitation for each month, each year
vw.temp.mth.yr <- vw_data %>% 
  group_by(Year, Month) %>%
  summarize(ave_temp=mean(Mean.Temp, na.rm=TRUE),
            ave_max_temp=mean(Max.Temp, na.rm=TRUE),
            ave_min_temp=mean(Min.Temp, na.rm=TRUE),
            max_temp=max(Max.Temp, na.rm=TRUE),
            min_temp=min(Min.Temp, na.rm=TRUE))

#is.nan(vw.temp.mth.yr$ave_max_temp)
#vw.temp.mth.yr %>% filter(max_temp==-Inf)
## need to clean up missing values
## even with na.rm used above, if there are only NA for a given month, 
##  it will return NaN for mean and Inf or -Inf for min/max
## clean up
vw.temp.mth.yr <- vw.temp.mth.yr %>% mutate(
  ave_temp=ifelse(is.nan(ave_temp),NA,ave_temp),
  ave_max_temp=ifelse(is.nan(ave_max_temp),NA,ave_max_temp),
  ave_min_temp=ifelse(is.nan(ave_min_temp),NA,ave_min_temp),
  max_temp=ifelse(max_temp==-Inf|max_temp==Inf,NA, max_temp),
  min_temp=ifelse(min_temp==-Inf|min_temp==Inf,NA, min_temp)
)

```

```{r MONTHLY TEMP COMBINED MONTHS ACROSS YEARS}
vw.temp.mth <- vw.temp.mth.yr %>% group_by(Month) %>%
  summarize(ave_temp=mean(ave_temp, na.rm=TRUE),
            ave_max_temp=mean(ave_max_temp, na.rm=TRUE),
            ave_min_temp=mean(ave_min_temp, na.rm=TRUE)) %>%
  mutate(pc_rank=percent_rank(ave_temp),
         rank=rank(desc(ave_temp)))

## filter for current year to overlay current year metrics on chart
vw.temp.mth.curr.yr <- vw.temp.mth.yr %>% filter(Year==year(Sys.Date()))
## need to have full 12 months for chart (even though some empty)
## get highest month with data in current year
fms <- max(as.numeric(vw.temp.mth.curr.yr$Month))+1
## create data frame for blank months
vw.temp.yr.future.mth <- data.frame(Year=year(Sys.Date()),
                                      Month=c(fms:12))
## convert Month field in current yr from factor to integer for binding
vw.temp.mth.curr.yr$Year <- as.integer(vw.temp.mth.curr.yr$Year)
vw.temp.mth.curr.yr$Month <- as.integer(vw.temp.mth.curr.yr$Month)

## add blank rows to current yr
vw.temp.mth.curr.yr <- bind_rows(vw.temp.mth.curr.yr,
                                    vw.temp.yr.future.mth) 

## Ave temp by month ####
chart.title <- "Average Temp. Month (blue is ave. min/max; red marker is curr. yr.)"
ggplot(vw.temp.mth, aes(x=Month, y=ave_temp))+geom_col()+
  geom_errorbar(aes(ymin=ave_min_temp, ymax=ave_max_temp), width=.2,
                 position=position_dodge(.9), color='blue')+
  geom_point(data=vw.temp.mth.curr.yr, aes(x=Month, y=ave_temp),shape=18, color='red', size=5)+
  scale_y_continuous(expand=c(0,0))+
  ggtitle(chart.title)+theme_classic()
## Ave precip by month in order
chart.title <- "Months Ranked by Ave. Temp"
ggplot(vw.temp.mth, aes(x=reorder(Month, -ave_temp), y=ave_temp))+geom_col()+
  ggtitle(chart.title)+theme_classic()

```

### Monthly Temperatures: Ranges

```{r TEMPERATURE RANGES}
## Distribution of temperature by month ####

## Range in temperature by month ####
chart.title <- "Monthly Temp. Ranges during Period (Red markers: current yr.)"
ggplot(vw.temp.mth.yr, aes(x=Month, y=ave_temp))+geom_boxplot()+
  geom_point(data=vw.temp.mth.curr.yr, aes(x=Month, y=ave_temp),shape=18, color='red', size=5)+
  ggtitle(chart.title)+theme_classic()
```

### Monthly Temperature Trends over Years

```{r MONTHLY TEMP TRENDS}
## annual trends in month-by-month temp
chart.title <- "Ave. Mthly Temps. All Years"
## need to convert Month to integer for line chart
vw.temp.mth.yr$Month <- as.integer(vw.temp.mth.yr$Month)
## set chart to variable for use with plotly
plot.mth.yr <- ggplot(vw.temp.mth.yr, aes(x=Month, y=ave_temp, color=Year))+geom_line()+scale_x_continuous(breaks=c(1:12))+
  theme_classic()+ggtitle(chart.title)
## use plotly for interaction such as highlight specific years
ggplotly(plot.mth.yr)

```

### Monthly Temp. Comparison YoY

For each month, what has been the pattern in ave. temperature over the years?

```{r PLOT MONTHS YOY TEMP, fig.width=5, fig.height=14}
## show each mth side-by-side YoY to see patterns, highs, lows

## facet_wrap is condensed but too many years to be able to read x axis
# ggplot(vw.precip.mth.yr, aes(x=Year, y=ttl_precip))+geom_bar(stat='identity')+
#   facet_wrap(Month~.)+
#   theme(axis.text.x  = element_text(angle=90, vjust=0.5))

## facet_grid is too wide and short unless re-sized
ggplot(vw.temp.mth.yr, aes(x=Year, y=ave_temp))+geom_col()+
  facet_grid(Month~.)+
  theme_bw()+
  theme(axis.text.x  = element_text(angle=90, vjust=0.5))
  
```

### Seasonal Temperatures

#### Season definitions:

* Winter: Dec, Jan, Feb
* Spring: Mar, Apr, May
* Summer: Jun, Jul, Aug
* Fall: Sep, Oct, Nov

*Note: the year for a winter season is applied to year at end of season. So winter from Dec 2018 to Feb 2019 is considered winter of 2019.*
